Sensitivity Analysis

### 2024-05-12 SFE_article_Sensitivity_Analysis_short - Chamomile/Article_Sensitivity_Analsysis_short-Chamomile.tex
12.05

The sensitivity analysis is a tool to understand the parametric dependence in the dynamic behaviour of the analysed system. The presented formulation involves derivative-based local sensitivity analysis of the model solution with respect to selected parameters and controls. The sensitivity equations can be obtained in various ways, but the automatic differentiation technique was implemented to obtain the sensitivity equations in this work. By identifying which parameters are influential at each period of the simulation, local sensitivity analysis can yield valuable information for guiding the process modelling, the design of experiments or the model reduction. The local sensitivity analysis techniques consider the sensitivities at only a small region of parameter space, and the conclusions derived from such an analysis are limited to local conditions unless the discussed system is a linear model.

In this study, the local sensitivity analysis method was introduced to explore the dynamics of the supercritical extraction system, which consists of a set of partial differential equations. The sensitivity equations evaluate the influence of control variables such as flow rate, pressure, and inlet temperature on the state-space consisting of the concentration of the solute in solid and liquid phases, the fluid's enthalpy density, system pressure, and extraction yield. Every change in the control variables leads to a state-space change, eventually affecting the process performance.

Similarities between Figures \ref{fig:Sensitivty_F_y}, \ref{fig:Sensitivty_P_y} and \ref{fig:Sensitivty_T_y} can observed, although the sources of disturbances are different. All the plots stay initially plat, and this is because the extraction kinetic is mainly controlled by the concentration gradient. The largest sensitivity deviations are observed in the second extraction phase, which is mainly controlled by the internal mass transfer. These deviations are strongly related to the fluid velocity and the correlations used to connect the extraction kinetic parameters. As correlations used in this work are linear functions of the fluid density( {\color{red} article 1}), Figures \ref{fig:Sensitivty_P_y} and \ref{fig:Sensitivty_T_y} are characterised by the opposite changes in the fluid density, which leads to the sensitivity plots with similar shape but opposite sign. In the third extraction phase, the concentration gradient and the internal diffusion coefficient decrease with the solute concentration in the solid phase( it is more difficult to extract the solute from the particle core than from the outer shell). It is important to note that the local-sensitivity methods obtain the discussed results, which can be obtained at different operating conditions.

>>>>

The sensitivity analysis is a tool to understand the parametric dependence in the dynamic behaviour of the analysed system. The presented formulation involves derivative-based local sensitivity analysis of the model solution with respect to selected parameters and controls. The sensitivity equations can be obtained in various ways, but the automatic differentiation technique was implemented to obtain the sensitivity equations in this work. By identifying which parameters are influential at each period of the simulation, local sensitivity analysis can bring valuable information about process modelling, design of experiments or model reduction. The local sensitivity analysis techniques consider only a small region of parameter space, and the conclusions derived from such an analysis are limited to local conditions unless the discussed system is a linear model.

In this study, the local sensitivity analysis method was introduced to explore the dynamics of the supercritical extraction system, which consists of a set of partial differential equations. The sensitivity analysis was used to evaluate the influence of pressure increase on the state-space consisting of the solutes concentration in solid and liquid phases, the fluid's enthalpy density, system pressure, and extraction yield. Every change in the control variables leads to a state-space change, eventually affecting the process performance.

At given operating conditions, the step change of pressure enhance the mass transfer, which leads to observation faster loss of solute by the solid particles and consequently to negative sensitivities. Analogous response is visible in the fluid phase, where the concentration of solute increase, which is visible as positive sensitivities. As more solute is transported to the fluid phase in shorter time, the extraction yield is also improved.

%Similarities between Figures \ref{fig:Sensitivty_F_y}, \ref{fig:Sensitivty_P_y} and \ref{fig:Sensitivty_T_y} can observed, although the sources of disturbances are different. All the plots stay initially plat, and this is because the extraction kinetic is mainly controlled by the concentration gradient. The largest sensitivity deviations are observed in the second extraction phase, which is mainly controlled by the internal mass transfer. These deviations are strongly related to the fluid velocity and the correlations used to connect the extraction kinetic parameters. As correlations used in this work are linear functions of the fluid density( {\color{red} article 1}), Figures \ref{fig:Sensitivty_P_y} and \ref{fig:Sensitivty_T_y} are characterised by the opposite changes in the fluid density, which leads to the sensitivity plots with similar shape but opposite sign. In the third extraction phase, the concentration gradient and the internal diffusion coefficient decrease with the solute concentration in the solid phase( it is more difficult to extract the solute from the particle core than from the outer shell). It is important to note that the local-sensitivity methods obtain the discussed results, which can be obtained at different operating conditions.

### 2024-05-12 SFE_article_Sensitivity_Analysis_short - Chamomile/Sections/Sensitivity_Pressure.tex
12.05

	Figure \ref{fig:Sensitivty_P_CF} shows the sensitivity of the solute concentration in the fluid phase as a response to the pressure change. As explained above, an increase in pressure enhanced the mass transfer, resulting in concentration front moving through the system. Initially, the system response is low due to low response in the solid phase. \todo{TODO}
	Following this, the formation of a trailing front characterized by decreasing sensitivities can be observed. This trailing front can be explained by the decrease in the concentration gradient as the solute is increasingly extracted and the available concentration within the solid phase diminishes. As seen in the figure, sensitivities are declining towards zero. This asymptotic trend suggests the driving force for the mass transfer—the concentration gradient—has been minimized, and changes in pressure no longer significantly impact the solute concentration in the fluid phase.

@@

	Figure \ref{fig:Sensitivty_P_y} illustrates the temporal evolution of the extraction yield's sensitivity to changes in pressure within a supercritical fluid extraction system.

	Initially, the curve displays an almost flat profile, suggesting a latency in the system's response to pressure changes. A minor initial decrease in sensitivity, indicated by a slight dip below zero, may arise due to a decrease in the velocity of the fluid phase. For a constant mass flow rate, an increase in fluid density due to heightened pressure results in a lower fluid velocity, potentially leading to this reduced sensitivity.

	As the process continues, the curve ascends sharply, indicating a positive yield response to pressure. This increase in sensitivity can be associated with the enhanced solubility and mass transfer capabilities imparted by higher pressures, improving the rate at which solute is extracted into the fluid phase. The peak in dy/dP reflects the most responsive period of the system, where pressure changes are highly effective in improving yield.

	Beyond the peak, the sensitivity declines, converging towards zero. This descending curve represents a depletion phase: as the available solute in the solid phase diminishes, the potential for further increasing the yield through pressure alone is reduced. The remaining solute concentration becomes a limiting factor, and the mass transfer driving force, initially augmented by pressure, no longer plays a dominant role.

>>>>

	Figure \ref{fig:Sensitivty_P_CF} shows the sensitivity of the solute concentration in the fluid phase with respect to the pressure change. As explained above, an increase in pressure enhances the mass transfer, resulting in the concentration front moving through the system. The system response is initially low, reflecting the idle period discussed above. Later, the sensitivities increased due to enhanced mass transfer. As the fluid phase is mobile, the advection affects the sensitivities, which causes the sensitivities to move along the system analogously to the solute in the fluid phase.

	After the growth phase, the sensitivities start to decrease, which can be explained by the decrease in the concentration gradient. As seen in Figure \ref{fig:Sensitivty_P_CF}, sensitivities decline towards zero. This asymptotic trend suggests the driving force for the mass transfer—the concentration gradient—has been minimized, and changes in pressure no longer significantly impact the solute concentration in the fluid phase.

@@

	Figure \ref{fig:Sensitivty_P_y} illustrates how sensitive the extraction yield is to pressure change. Initially, the sensitivity curve stays almost flat, suggesting a latency in the system's response to pressure changes. A minor decrease in sensitivity arises due to a decrease in the velocity of the fluid phase. As the fluid moves slower, it reaches the end of the extractor later, which causes negative sensitivities.

	As the process continues and the solute reaches the extractor's outlet, the sensitivity curve increases rapidly, indicating an improvement in the process efficiency. This increase in sensitivity can be associated with the enhanced solubility and mass transfer imparted by higher pressures. The peak in $\frac{dy}{dP}$ reflects when the deviation from the original system is the largest.

	Beyond the peak, the sensitivity declines, converging towards zero. This descending curve represents a depletion phase: as the available solute in the solid phase diminishes, the potential for further increasing the yield through pressure alone is reduced. The remaining concentration gradient becomes a limiting factor, and the enhanced mass transfer no longer plays a dominant role compared to the original system.
